Loblaws Seasonal Discount Patterns: Peak Sales Frequency in Fall and Deeper Discounts in Early Spring

### 2024-11-21 paper/paper.qmd
committing formatting changes to paper.qmd

## Monthly Frequency of Items on Sale

@fig-monthly-sales reveals significant trends in the volume of sales across different months. The chart demonstrates a steep increase in the number of items on sale starting in June, reaching a peak in October, followed by a sharp drop in November. This pattern suggests that Loblaws ramps up sales activity in the late summer and early fall, likely aiming to clear inventory before the holiday season. The high frequency of items on sale in October could be associated with seasonal transitions or targeted discounting to attract customers before the end-of-year shopping period. The dramatic decline in November may indicate a pause before holiday promotions, aligning with common retail strategies to manage demand.

```{r}
#| echo: false
#| eval: true
#| label: fig-monthly-sales
#| fig-cap: "Monthly trend in the frequency of items on sale at Loblaws, showing peak sale activity in October and a drop in November."
#| warning: false
#| message: false

@@

```

@@

## Density of Discount Percentages by Season

@fig-discount-density-season illustrates that winter tends to have smaller, consistent discounts, while fall shows a broader spread, including more substantial discounts. Spring and summer exhibit moderate discount distributions. This seasonal comparison underscores the strategic use of varied discount levels by Loblaws, with deeper discounts appearing more frequently in fall and winter, possibly to drive purchases during key shopping periods.
```{r}
#| echo: false
#| eval: true
#| label: fig-discount-density-season
#| fig-cap: "Density plot of discount percentages by season, highlighting fall as the season with a broader range of discounts, including higher values."
#| warning: false

@@

Collectively, the visualizations provide insights into Loblaws' seasonal and monthly discount strategies. The data suggests that Loblaws increases the frequency of sales in the latter half of the year while moderating the average discount depth, likely balancing consumer demand with profit margins. The fall season appears to be a critical period for aggressive discounting, likely tied to seasonal shopping behaviors. By analyzing both the frequency and depth of discounts, this study highlights the complexity of retail discount strategies and provides a basis for further investigation into how these strategies impact consumer behavior and sales outcomes.

>>>>

## Monthly Frequency of Items on Sale
```{r}
#| echo: false
#| eval: true
#| label: fig-monthly-sales
#| fig-cap: "Monthly trend in the frequency of items on sale at Loblaws, showing peak sale activity in October and a drop in November."
#| warning: false
#| message: false

@@

```
@fig-monthly-sales reveals significant trends in the volume of sales across different months. The chart demonstrates a steep increase in the number of items on sale starting in June, reaching a peak in October, followed by a sharp drop in November. This pattern suggests that Loblaws ramps up sales activity in the late summer and early fall, likely aiming to clear inventory before the holiday season. The high frequency of items on sale in October could be associated with seasonal transitions or targeted discounting to attract customers before the end-of-year shopping period. The dramatic decline in November may indicate a pause before holiday promotions, aligning with common retail strategies to manage demand.

@@

## Density of Discount Percentages by Season
```{r}
#| echo: false
#| eval: true
#| label: fig-discount-density-season
#| fig-cap: "Density plot of discount percentages by season, highlighting fall as the season with a broader range of discounts, including higher values."
#| warning: false

@@

@fig-discount-density-season illustrates that winter tends to have smaller, consistent discounts, while fall shows a broader spread, including more substantial discounts. Spring and summer exhibit moderate discount distributions. This seasonal comparison underscores the strategic use of varied discount levels by Loblaws, with deeper discounts appearing more frequently in fall and winter, possibly to drive purchases during key shopping periods. Collectively, the visualizations provide insights into Loblaws' seasonal and monthly discount strategies. The data suggests that Loblaws increases the frequency of sales in the latter half of the year while moderating the average discount depth, likely balancing consumer demand with profit margins. The fall season appears to be a critical period for aggressive discounting, likely tied to seasonal shopping behaviors. By analyzing both the frequency and depth of discounts, this study highlights the complexity of retail discount strategies and provides a basis for further investigation into how these strategies impact consumer behavior and sales outcomes.
